---
title: "My R code from the Level 1 and 2 exercises of Kaggle's Learn Machine Learning series"
author: "Michael Skolnik"
subtitle: ""
date: "2018-04-12"
slug: "home-prices-r"
tags: ["R", "machine learning"]
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## *Learn Maching Learning* series on Kaggle in R

This is my R code for the *Learn Machine Learning* series on Kaggle. I've already done the Python one, which is on Kaggle located [here](https://www.kaggle.com/learn/machine-learning). The data used is from the [*Home Prices: Advanced Regression Techniques*](https://www.kaggle.com/c/house-prices-advanced-regression-techniques) competition.

I'm going to run the completed code from level 1 first and then run each level 2 section.

```{r}
if (!require("randomForest")) {
  install.packages("randomForest", repos="http://cran.rstudio.com/")
  library(randomForest)
}

if (!require("dplyr")) {
  install.packages("dplyr", repos="http://cran.rstudio.com/")
  library(dplyr)
}

if (!require("caTools")) {
  install.packages("dplyr", repos="http://cran.rstudio.com/")
  library(caTools)
}

# Save filepath to variable
training_data_filepath <- "C:/Development/Kaggle/House Prices - Advanced Regression Techniques/train.csv"

# Import data
dataset <- read.csv(training_data_filepath)

# Split data into training and validation data, for both predictors and target.
set.seed(42)
split <- sample.split(dataset, SplitRatio=0.7)  # for training data
training_set <- subset(dataset, split==TRUE)
test_set <- subset(dataset, split==FALSE)

# Create the training and tests dataframe with only numeric predictors
nums <- unlist(lapply(training_set, is.numeric))
training_set <- training_set[, nums]
test_set <- test_set[, nums]

# Drop NAs
# training_set <- training_set[complete.cases(training_set), ]
# test_set <- test_set[complete.cases(test_set), ]

# Replace NA with 0
# training_set[is.na(training_set)] <- 0

# Create the predictor variable
X <- subset(training_set, select = -c(Id, SalePrice))

# Select the target variable and call it y
y <- training_set$SalePrice

# Fitting Random Forest Regression to the dataset
regressor <- randomForest(x=X, y=y, ntree=100)
```

If I run this I get an error because of the NA values in the data. I need to impute the missing values first.

```{r}

```
